---
title: "Ustedeo and voseo in Dataland"
subtitle: "Study presentation"
author: "Andrea V Carbone"
institute: "Rutgers"
date: "2025/5/5"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"  # Adjust slide size
      slideNumberFormat: "%current%"  
      css: ["default", "custom.css"]
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Load the necessary libraries
library(tidyverse)
library(kableExtra) # for nicer table output

# Load the raw data
raw_data <- read_csv("../data/raw/binary_outcome_data.csv")


```

#slide 2 tidying data

This is the raw data:
```{r echo=FALSE, warning=FALSE, message=FALSE}
# Inspect the first few rows of the data
head(raw_data)
```
And this is the tidied data:
```{r echo=FALSE, warning=FALSE, message=FALSE}
# Tidy the data
tidied_data <- raw_data %>%
  rename(subject = participant_id) %>% # Rename participant_id to subject
  separate(item, into = c("setting", "question"), sep = "_", extra = "merge") %>%
  mutate(question = paste0("q", str_remove(question, "^[a-z]+"))) %>%
  mutate(question = as.factor(question)) %>%
  select(-response) %>%
  rename(response = response_label) %>% # Rename response_label to response
  mutate(response = as.factor(response),
         setting = as.factor(setting),
         subject = as.factor(subject)) 

# Inspect the tidied data
head(tidied_data)
glimpse(tidied_data)

# Save the tidied data
write_csv(tidied_data, "../data/tidy/tidied_data.csv")
```
---
# Step 2: Build the logistic regression models using glm()

# Model 1: Effect of Setting
``` {r echo=TRUE, warning=FALSE, message=FALSE}
model1 <- glm(I(response == "usted") ~ setting,
              family = binomial(link = "logit"),
              data = tidied_data)

```
??
This code builds a logistic regression model to examine if there is a relationship between the *setting and the probability of a speaker choosing 'usted'. The model will estimate how the odds of choosing 'usted' change depending on the setting, and it will also tell us if this relationship is statistically significant based on the data.


# Step 2: Build the logistic regression models using glm()

# Model 2: Effect of Age
``` {r echo=TRUE, warning=FALSE, message=FALSE}
model2 <- glm(I(response == "usted") ~ age,
              family = binomial(link = "logit"),
              data = tidied_data)
```
??
This code builds a logistic regression model to examine if there is a relationship between the speaker's *age* and the probability of choosing 'usted'. The model will estimate how the odds of choosing 'usted' change with the speaker's age, and it will also tell us if this relationship is statistically significant based on the data.

# Step 2: Build the logistic regression models using glm()

# Model 3: Additive Effects of Setting and Age
``` {r echo=TRUE, warning=FALSE, message=FALSE}
model3 <- glm(I(response == "usted") ~ setting + age,
              family = binomial(link = "logit"),
              data = tidied_data)
```
??
This code builds a logistic regression model to examine the relationships between both the setting and the speaker's age and the probability of choosing 'usted'. The model will estimate how the odds of choosing 'usted' change depending on the setting, and also how they change with the speaker's age, assuming these two factors have independent effects. It will also tell us if each of these relationships is statistically significant based on the data.The summary will tell us the relation between *age* end the choice of "usted" while holding the setting constant. Similarl ywill tell us about the relationship between *setting* and the choice of "usted" while holding age constant.

# Step 2: Build the logistic regression models using glm()
#Model 4: Interaction Effect of Setting and Age
``` {r echo=TRUE, warning=FALSE, message=FALSE}
model4 <- glm(I(response == "usted") ~ setting * age,
              family = binomial(link = "logit"),
              data = tidied_data)
```
??
This is the most interesting model. It tests if the effect of setting depends on the age of the speaker. For example, maybe the setting has a strong effect on younger speakers but a weaker effect on older speakers.

This code builds a logistic regression model to examine how the *setting* and the speaker's *age* together influence the probability of choosing 'usted'. Unlike Model 3, this model also considers if the effect of the setting (formal vs. informal) on the choice of 'usted' is different for speakers of different ages. The model will estimate the individual effects of *setting* and *age*, as well as how their combined effect might change the odds of choosing 'usted'. It will also tell us if these individual and combined relationships are statistically significant based on the data. A small p-value (typically less than 0.05) suggests that the predictor (el tipo de valores que elegimos para el eje x) is statistically significant.
---
``` {r echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(kableExtra)

options(knitr.kable.NA = "", knitr.table.format = "html")

kable_styling_global <- function(kable_input) {
  kable_input %>%
    kable_styling(font_size = 10, bootstrap_options = c("striped", "hover"), full_width = FALSE)
}


```


``` {r echo=FALSE, warning=FALSE, message=FALSE}
summary_model1_df <- as.data.frame(summary(model1)$coefficients)


colnames(summary_model1_df) <- gsub("Pr\\(>\\|z\\|\\)", "p-value", colnames(summary_model1_df))

kable(summary_model1_df, caption = "Summary of Model 1 (Setting)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)%>%
  kable_styling_global() %>%
  column_spec(1:ncol(summary_model1_df), width = "2cm")
```




``` {r echo=FALSE, warning=FALSE, message=FALSE}
summary_model2_df <- as.data.frame(summary(model2)$coefficients)
colnames(summary_model2_df) <- gsub("Pr\\(>\\|z\\|\\)", "p-value", colnames(summary_model2_df))
kable(summary_model2_df, caption = "Summary of Logistic Regression Model 2 (Age)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)%>%
  kable_styling_global() %>%
  column_spec(1:ncol(summary_model2_df), width = "2.5cm")
```


``` {r echo=FALSE, warning=FALSE, message=FALSE}
summary_model3_df <- as.data.frame(summary(model3)$coefficients)
colnames(summary_model3_df) <- gsub("Pr\\(>\\|z\\|\\)", "p-value", colnames(summary_model3_df))
kable(summary_model3_df, caption = "Summary of Logistic Regression Model 3 (Setting + Age)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  column_spec(1:ncol(summary_model3_df), width = "2.5cm") 
```
``` {r echo=FALSE, warning=FALSE, message=FALSE}
summary_model4_df <- as.data.frame(summary(model4)$coefficients)
colnames(summary_model4_df) <- gsub("Pr\\(>\\|z\\|\\)", "p-value", colnames(summary_model4_df))
kable(summary_model4_df, caption = "Summary of Logistic Regression Model 4 (Setting * Age)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  column_spec(1:ncol(summary_model4_df), width = "2.5cm")
```
***Setting interacting with Age: 1 settinginformal: 
1 'settinginformal': The estimate of 1.130 tells us the difference in the odds of choosing 'usted' between the informal and formal settings when age is zero. 
2 'age': The estimate of -0.026 tells us the effect of age on the odds of choosing 'usted' in the formal setting.  
3 'settinginformal:age': The estimate of -0.012 for the interaction term indicates that the effect of age on the log-odds of choosing 'usted' differs between the informal and formal settings. Because it's negative, and the 'age' estimate is also negative, the negative effect of age is stronger in the informal setting.  

---
#"Model 1: Effect of Setting"

```{r echo=FALSE, warnings=FALSE}
# 1. Create prediction data
predicted_model1 <- tidied_data %>%
  distinct(setting) %>%
  mutate(probability = predict(model1, newdata = ., type = "response"))

# 2. Create the plot
plot_model1 <- ggplot(predicted_model1, aes(x = setting, y = probability)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Predicted Probability of Choosing 'Usted' by Setting (Model 1)",
    x = "Setting",
    y = "Predicted Probability"
  ) +
  theme_minimal() +
  geom_text(aes(label = round(probability, 2)), vjust = -0.3)

# Print the plot
print(plot_model1)
```

---
# "Model 2: Effect of Age"

```{r echo=FALSE, warnings=FALSE}

# 1. Create prediction data
age_range <- range(tidied_data$age)
predicted_model2 <- data.frame(age = seq(age_range[1], age_range[2], length.out = 100))
predicted_model2$probability <- predict(model2, newdata = predicted_model2, type = "response")

# 2. Create the plot
plot_model2 <- ggplot(predicted_model2, aes(x = age, y = probability)) +
  geom_line(color = "darkgreen", linewidth = 1) + # Changed size to linewidth
  labs(
    title = "Predicted Probability of Choosing 'Usted' by Age (Model 2)",
    x = "Age",
    y = "Predicted Probability"
  ) +
  theme_minimal()

# Print the plot
print(plot_model2) # Added print()
```

---
##  "Model 3: Additive Effects of Setting and Age"

```{r echo=FALSE, warnings=FALSE}
# 1. Create prediction data
age_range <- range(tidied_data$age)
predicted_model3 <- expand.grid(
  setting = unique(tidied_data$setting),
  age = seq(age_range[1], age_range[2], length.out = 100)
)
predicted_model3$probability <- predict(model3, newdata = predicted_model3, type = "response")

# 2. Create the plot
plot_model3 <- ggplot(predicted_model3, aes(x = age, y = probability, color = setting)) +
  geom_line(linewidth = 1) + # Changed size to linewidth
  labs(
    title = "Predicted Probability of Choosing 'Usted' by Age and Setting (Model 3)",
    x = "Age",
    y = "Predicted Probability",
    color = "Setting"
  ) +
  theme_minimal()

# Print the plot
print(plot_model3) # Added print()
```

---
##  "Model 4: Interaction Effect of Setting and Age"

```{r echo=FALSE, warnings=FALSE}
# 1. Create prediction data
age_range <- range(tidied_data$age)
predicted_model4 <- expand.grid(
  setting = unique(tidied_data$setting),
  age = seq(age_range[1], age_range[2], length.out = 100)
)
predicted_model4$probability <- predict(model4, newdata = predicted_model4, type = "response")

# 2. Create the plot
plot_model4 <- ggplot(predicted_model4, aes(x = age, y = probability, color = setting)) +
  geom_line(linewidth = 1) + # Changed size to linewidth
  labs(
    title = "Predicted Probability of Choosing 'Usted' by Age and Setting (Model 4)",
    x = "Age",
    y = "Predicted Probability",
    color = "Setting"
  ) +
  theme_minimal()

# Print the plot
print(plot_model4) # Added print()
```
---
Thank you

