---
title: "Code"
output: html_document
date: "2025-05-12"
---





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```
```{r echo=FALSE, include=FALSE, load-and-tidy-data}
# Load raw data
raw_data <- read_csv("../data/raw/binary_outcome_data.csv")
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
# Inspect the first few rows of the data
library(flextable)
library(magrittr)  # Ensures %>% is available

head(raw_data) %>%  
  flextable() %>% 
  set_caption(caption = "Sample of Raw Data") %>% 
  autofit()

```




```{r echo=FALSE, warning=FALSE, message=FALSE}
# Tidy the data
tidied_data <- raw_data %>%
  rename(subject = participant_id) %>% # Rename participant_id to subject
  separate(item, into = c("setting", "question"), sep = "_", extra = "merge") %>%
  mutate(question = paste0("q", str_remove(question, "^[a-z]+"))) %>%
  mutate(question = as.factor(question)) %>%
  select(-response) %>%
  rename(response = response_label) %>% # Rename response_label to response
  mutate(response = as.factor(response),
         setting = as.factor(setting),
         subject = as.factor(subject)) 

head(tidied_data) %>%  
  flextable() %>% 
  set_caption(caption = "Sample of Tidied Data") %>% 
  autofit()

```




# Model Comparison
## Model 1: Effect of Setting
``` {r echo=TRUE, warning=FALSE, message=FALSE}
model1 <- glm(I(response == "usted") ~ setting,
              family = binomial(link = "logit"),
              data = tidied_data)

```

## Model 2: Effect of Age
``` {r echo=TRUE, warning=FALSE, message=FALSE}
model2 <- glm(I(response == "usted") ~ age,
              family = binomial(link = "logit"),
              data = tidied_data)
```

## Model 3: Additive Effects of Setting and Age
``` {r echo=TRUE, warning=FALSE, message=FALSE}
model3 <- glm(I(response == "usted") ~ setting + age,
              family = binomial(link = "logit"),
              data = tidied_data)
```

## Model 4: Interaction Effect of Setting and Age
``` {r echo=TRUE, warning=FALSE, message=FALSE}
model4 <- glm(I(response == "usted") ~ setting * age,
              family = binomial(link = "logit"),
              data = tidied_data)
```
# Model Results
``` {r echo=FALSE, warning=FALSE, message=FALSE}
summary_model1_df <- as.data.frame(summary(model1)$coefficients)


colnames(summary_model1_df) <- gsub("Pr\\(>\\|z\\|\\)", "p-value", colnames(summary_model1_df))
knitr::kable(summary_model1_df, caption = "Summary of Model 1 (Setting)")
```
The model in Table 3 shows that the odds of using "usted" are significantly lower in informal settings compared to formal ones. Specifically, being in an informal setting reduces the odds of using "usted" (estimate = -0.67, p < 0.001). The intercept suggests the baseline odds of using "usted" in formal settings.
Figure 1 displays the predicted probability of ‘usted’ use as a function of setting, based on Model 1.

```{r echo=FALSE, warnings=FALSE}

# 1. Create prediction data
predicted_model1 <- tidied_data %>%
  distinct(setting) %>%
  mutate(probability = predict(model1, newdata = ., type = "response"))

# 2. Create the plot
plot_model1 <- ggplot(predicted_model1, aes(x = setting, y = probability)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Predicted Probability of Choosing 'Usted' by Setting",
    caption= "Figure 1",
    x = "Setting",
    y = "Predicted probability (“usted”)"
  ) +
  theme_minimal() +
  geom_text(aes(label = round(probability, 2)), vjust = -0.3)

# Print the plot
print(plot_model1)
```

---

``` {r echo=FALSE, warning=FALSE, message=FALSE}
summary_model2_df <- as.data.frame(summary(model2)$coefficients)
colnames(summary_model2_df) <- gsub("Pr\\(>\\|z\\|\\)", "p-value", colnames(summary_model2_df))
knitr::kable(summary_model2_df, caption = "Summary of Logistic Regression Model 2 (Age)")
```
Model 2, in Table 4, shows that age is positively associated with the use of "usted". For each additional year of age, the odds of using "usted" increase (estimate = 0.031, p < 0.001). Older speakers are thus more likely to use "usted" than younger ones.
Figure 2 shows the predicted probability of ‘usted’ use across the observed range of speaker ages.
```{r echo=FALSE, warnings=FALSE}
# 1. Create prediction data
age_range <- range(tidied_data$age)
predicted_model2 <- data.frame(age = seq(age_range[1], age_range[2], length.out = 100))
predicted_model2$probability <- predict(model2, newdata = predicted_model2, type = "response")

# 2. Create the plot
plot_model2 <- ggplot(predicted_model2, aes(x = age, y = probability)) +
  geom_line(color = "darkgreen", linewidth = 1) + # Changed size to linewidth
  labs(
    title = "Predicted Probability of Choosing 'Usted' by Age",
    caption= "Figure 2",
    x = "Age",
    y = "Predicted Probability (“usted”)"
  ) +
  theme_minimal()

# Print the plot
print(plot_model2) # Added print()
```
---

``` {r echo=FALSE, warning=FALSE, message=FALSE}
summary_model3_df <- as.data.frame(summary(model3)$coefficients)
colnames(summary_model3_df) <- gsub("Pr\\(>\\|z\\|\\)", "p-value", colnames(summary_model3_df))
knitr::kable(summary_model3_df, caption = "Summary of Logistic Regression Model 3 (Setting + Age)")
```
The findings in Model 1 and Model 2 are confirmed in Model 3, as seen in Table 5.Informal settings still significantly reduce the odds of using "usted" (estimate = -0.71, p < 0.001), while age continues to increase the odds (estimate = 0.032, p < 0.001).
Both setting and age independently contribute to explaining variation in "usted" use.
Figure 3 presents the additive effects of setting and age on the probability of ‘usted’ use.
```{r echo=FALSE, warnings=FALSE}
# 1. Create prediction data
age_range <- range(tidied_data$age)
predicted_model3 <- expand.grid(
  setting = unique(tidied_data$setting),
  age = seq(age_range[1], age_range[2], length.out = 100)
)
predicted_model3$probability <- predict(model3, newdata = predicted_model3, type = "response")

# 2. Create the plot
plot_model3 <- ggplot(predicted_model3, aes(x = age, y = probability, color = setting)) +
  geom_line(linewidth = 1) + # Changed size to linewidth
  labs(
    title = "Pred. Probability of Choosing 'Usted' by Age and Setting",
    caption = "Figure 3",
    x = "Age",
    y = "Predicted Probability (“usted”)",
    color = "Setting"
  ) +
  theme_minimal()

# Print the plot
print(plot_model3) # Added print()
```



``` {r echo=FALSE, warning=FALSE, message=FALSE}
summary_model4_df <- as.data.frame(summary(model4)$coefficients)
colnames(summary_model4_df) <- gsub("Pr\\(>\\|z\\|\\)", "p-value", colnames(summary_model4_df))
knitr::kable(summary_model4_df, caption = "Summary of Logistic Regression Model 4 (Setting * Age)")
```
  
The model in Table 6 tests whether the effect of age depends on setting.
    • The main effects remain: informal settings reduce "usted" use, and age increases it.
    • The interaction term (settinginformal:age) is not statistically significant (p = 0.17), meaning there's no strong evidence that the effect of age differs between formal and informal settings.
Figure 4 visualizes the interaction between setting and age, showing predicted probabilities of ‘usted’ use for different combinations of these predictors.
```{r echo=FALSE, warnings=FALSE}
# 1. Create prediction data
age_range <- range(tidied_data$age)
predicted_model4 <- expand.grid(
  setting = unique(tidied_data$setting),
  age = seq(age_range[1], age_range[2], length.out = 100)
)
predicted_model4$probability <- predict(model4, newdata = predicted_model4, type = "response")

# 2. Create the plot
plot_model4 <- ggplot(predicted_model4, aes(x = age, y = probability, color = setting)) +
  geom_line(linewidth = 1) + # Changed size to linewidth
  labs(
    title = "Pred. Probability of Choosing 'Usted' by Age and Setting",
    caption = "Figure 4",
    x = "Age",
    y = "Predicted Probability (“usted”)",
    color = "Setting"
  ) +
  theme_minimal()

# Print the plot
print(plot_model4) # Added print()
```

